---
title: "Untitled"
author: "Pushkar Sinha"
date: "7/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## Fit a linear regression model
model = lm(sales ~ adv)
summary(model)

```


Coefficients:
              Estimate Std. Error   t value    Pr(>|t|)    
(Intercept)  -157.3301   145.1912    -1.084        0.29    
adv             2.7721     0.2794     9.921    8.87e-10 
Residual standard error: 101.4 on 23 degrees of freedom

The estimate for b1 is 2.7721. The variance estimate is 0.27942. The sampling distribution is a t-distribution with 23 degrees of freedom.
The estimate for b0 is -157.3301. The variance estimate is 145.19122.
The estimate for b1 is statistically significant, as evidenced by a p-value of 8.87×10-10


```{r}
tvalue = 9.921
##H0: b1 = 0 ; Ha: b1 != 0
2 * pt(tvalue, 23, lower.tail = FALSE)
2 * (1- pt(tvalue, 23))

##H0: b1 = 1 ; Ha: b1 != 1
t = (2.7721 - 1)/0.2794
2 * pt(t, 23, lower.tail = FALSE)
2 * (1- pt(t, 23))

## Statistically Positive
##H0: b1 <= 0 ; Ha - b1 > 0
pt(tvalue, 23, lower.tail = FALSE)
1- pt(tvalue, 23)

## Statistically Negative
##H0: b1 >= 0 ; Ha - b1 < 0
pt(tvalue, 23, lower.tail = TRUE)
pt(tvalue, 23)
```

## Confidence Interval for Coefficients

```{r}
confint(model, level=0.99)
confint(model, level=0.95)

```
                   0.5 %         99.5 %
(Intercept)  -564.930546      250.27032
adv             1.987712        3.55652


## Prediction
```{r}
new = data.frame(adv = 300)
predict.lm(model, new, interval = "predict", level = 0.99)
predict.lm(model, new, interval = "predict", level = 0.95)
predict.lm(model, new, interval = "confidence", level = 0.99)
predict.lm(model, new, interval = "confidence", level = 0.95)
```

       fit         lwr        upr
1 674.3047     338.712   1009.897

       fit         lwr        upr
1 674.3047    427.0146    921.5948

       fit         lwr        upr
1 674.3047    496.6497   851.9596

       fit        lwr         upr
1 674.3047    543.395     805.2143


## ##############################################################################################
## Confidence bands
## ##############################################################################################
```{r}
regplot.confbands.fun(Inflation.difference,Exchange.rate.change)
```


## ##############################################################################################
## Corr Matrix
## ##############################################################################################
```{r}
cor(buch,bush)
```

## ##############################################################################################
## Obtain estimated coefficients
## ##############################################################################################
```{r}
betas = coef(model)
```

## ##############################################################################################
## Linearity/Mean Zero, Independence and Constant Variance - X vs Y
## ##############################################################################################
```{r}
# SLR
plot(adv, sales)

ggplot(data=data, aes(x=chmax, y=performance)) +
  geom_point(alpha=I(0.2),color='blue') +
  xlab('Maximum Channels') +
  ylab('CPU Performance') +
  ggtitle('Maximum Channels vs CPU Performance') +
  geom_smooth(method= "lm",color='gray', se=FALSE)

```

## ##############################################################################################
## Linearity - scatterplots of the standardized residuals of model2 versus each quantitative predictor (X vs residuals)
## ##############################################################################################
```{r}
#SLR
plot(model$fitted, model$residuals)

# MLR
resids =rstandard(model2)
par(mfrow=c(2,3))
for (i in c(3:7)){
  col_name = names(fish2[i])
  plot(fish2[,i], resids, xlab= col_name, ylab = "S. Residuals")
  abline(h=0, col="red")
  lines(lowess(fish2[,i], resids), col='blue')
}
```

# ###############################################################################################
# Independence (Uncorrelated errors) and Constant Variance (fitted vs residuals)
# ###############################################################################################
```{r}
ggplot(data=data, aes(x=model1$fitted.values, y=model1$residuals)) +
  geom_point(alpha=I(0.4),color='darkorange') +
  xlab('Fitted Values') +
  ylab('Residuals') +
  ggtitle('Residual Plot') +
  geom_hline(yintercept=0)

# Plot of std. residuals versus fitted values
plot(model2$fitted.values, resids, xlab="Fitted Values", ylab=" S. Residuals")
lines(lowess(model2$fitted.values, resids), col='blue')
abline(h=0, col="red")
```

## ##############################################################################################
## Normality - Histogram and q-q plot of teh residuals
## ##############################################################################################

```{r}
hp = qplot(model1$residuals,
           geom="histogram",
           bins=11,
           main = "Histogram of Residuals",
           xlab = "Residuals",
           ylab = "Count",
           fill=I("blue"),
           alpha=I(0.2))
qqp = ggplot(data, aes(sample=model1$residuals)) +
stat_qq(alpha=I(0.2),color='darkorange') +
  stat_qq_line() +
  ggtitle("Q-Q Plot of Residuals")
ggarrange(hp, qqp, ncol=2, nrow=1)

## QQ plot
qqPlot(model1$residuals)

resids =rstandard(model2)
qqPlot(resids)

## Histogram Plots for normality
resids =rstandard(model2)
hist(resids, col="orange", nclass=15)

```

## ##############################################################################################
## Model Assumptions - One Variable 
## ##############################################################################################
```{r}
# Get standardized residuals
resids2 =rstandard(model3)
# Residual plots
par(mfrow=c(2,2))
plot(fish2$Total.Length, resids2 , main="Lineartiy/Mean Zero",
     xlab="Total.Length", ylab="Residuals")
abline(h=0, col="red")
lines(lowess(fish2$Total.Length, resids2), col='blue')
plot(model3$fitted.values, resids2 , main="Constant Variance/Correlation",
     xlab="Fitted Values", ylab="Residuals")
abline(h=0, col="red")
lines(lowess(model3$fitted.values, resids2), col='blue')
# Normality
hist(resids2 , col="orange", nclass=15)
qqPlot(resids2)

```

## ##############################################################################################
## Modeling  - SLR - Log transform
## ##############################################################################################
```{r}
model2 = lm(log(performance) ~ log(chmax + 1), data=data)

sp2 = ggplot(data=data, aes(x=log(chmax+1), y=log(performance))) +
  geom_point(alpha=I(0.2),color='blue') +
  xlab('Maximum Channels') +
  ylab('Performance') +
  ggtitle('Max Channels vs Performance') +
  geom_line(data=data, aes(x=log(chmax+1), y=model2$fitted.values))
rp2 = ggplot(data=data, aes(x=model2$fitted.values, y=model2$residuals)) +
  geom_point(alpha=I(0.4),color='darkorange') +
  xlab('Fitted Values') +
  ylab('Residuals') +
  ggtitle('Residual Plot') +
  geom_hline(yintercept=0)
hp2 = qplot(model2$residuals,
            geom="histogram",
            bins=11,
            main = "Histogram of Residuals",
            xlab = "Residuals",
            ylab = "Count",
            fill=I("blue"),
            alpha=I(0.2))
qqp2 = ggplot(data, aes(sample=model2$residuals)) +
  stat_qq(alpha=I(0.2),color='darkorange') +
  stat_qq_line() +
  ggtitle("Q-Q Plot of Residuals")
ggarrange(sp2, rp2, hp2, qqp2, ncol=2, nrow=2)

```




